---
title: "Perform a statistical adjustement (reweighting)"
author: "Dominique Emmanuel"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Theory

Let $X^1,\ldots, X^n$ be some categorical variables, and for each variable $X^i$ let $m^{i1},\ldots,m^{in_i}$ be its levels, and let $X^{i1},\ldots,X^{in_i}$ be the associated dummies. Let $w$ be an initial weight.
We call **adjustement** a weight $w'$ such that :

* $w'$ is as close as possible to $w$ (L2 norm)

* $\sum_{k=1}^Nw'_k=\sum_{k=1}^Nw_k$

* $\forall k~~0\leq w_{\text{min}} \leq w'_k \leq w_{\text{max}}$

* the weighted values of each level $m^{ij}$ is equal (and/or greater and/or lower) to specifed value $v^{ij}$.

Consequently

$$w' = \underset{
\substack{\forall i,j~~\sum_{k=1}^Nx_kX_k^{ij}=v^{ij}\\\sum_{k=1}^Nx_k=\sum_{k=1}^Nw_k\\\forall k~~w_{\text{min}} \leq x_k \leq w_{\text{max}}}}{\arg\min}~\left\|x-w\right\|^2$$

NB: Any equality $\sum_{k=1}^Nx_kX_k^{ij}=v^{ij}$ can be replaced by an inequality.

## Practice
The function `adjustement` allows to perform this optimisation.

Let's take a subest of `esoph` of 50 individuals :
```{r}
set.seed(123)
data <- esoph[sample(seq(nrow(esoph)),50), ]
w_initial <- rep(nrow(esoph)/nrow(data), nrow(data))
````


Let's define somme margins:
```{r}
table(esoph$agegp)/nrow(esoph)
table(esoph$alcgp)/nrow(esoph)
margins <- list(
  list(var_name = "agegp",
       value = c("25-34" = 0.17, "35-44" = 0.17, "45-54" = 0.17, "55-64" = 0.17),
       min = c("65-74" = 0.17, "75+" = 0.12)
  ),
  list(var_name = "alcgp",
       value = c("0-39g/day" = 0.26), 
       min = c("40-79" = 0.3),
       max = c("80-119" = 0.2)
  )
)
````
Let's perform the adjustement:
```{r}
library(adjustment)
adj <- adjustment(data = data, margins = margins, weight = w_initial, weight_min = 0.1, weight_max = 30)
adj$IsError
w <- adj$w
````

Let's verify:
```{r}
library(dplyr)
data$w <- w
data %>%
  group_by(agegp) %>%
  summarise(n = sum(w)) %>%
  merge(data %>% summarise(n0 = sum(w))) %>%
  mutate(mean = n/n0)

data %>%
  group_by(alcgp) %>%
  summarise(n = sum(w)) %>%
  merge(data %>% summarise(n0 = sum(w))) %>%
  mutate(mean = n/n0)

hist(w, 10, col = "steelblue", xlab = "new weight",border="white")
````

